Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-03 PandaEvolution.tex
minor corrections and comments in Section 3.

Titan's architecture, configuration and policies poses several challenges to the
integration with PanDA. The deployment model of PanDA Pilot is unfeasible on
Titan: PanDA Pilot requires to contact the Job Dispatcher of the PanDA Server to
pull jobs to execute but this is not possible on Titan because worker nodes do
not offer outbound network connectivity. Further, Titan does not support PanDA's
security model based on certificates and virtual organizations, making the
PanDA's approach to identity management also unfeasible. While DTNs offer wide
area network data transfer, an integration with ATLAS DDM is beyond the
functional and administrative scope of the current prototyping phase. Finally,
the specific characteristics of the execution environment, especially the
absence of local storage on the worker nodes and modules tailored to Compute
Node Linux, require reengineering of ATLAS application frameworks.

Currently, very few HEP applications can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows can
be offloaded from the Grid to Titan's large amount of cores. Further, when HEP
tasks can be partitioned to independent jobs, Titan worker nodes can be used to
execute up to 16 concurrent payloads, one for each available core. Given these
constraints and challenges, the type of task most suitable for execution at the
moment on Titan is Monte Carlo detector simulation. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and with
small input data requirements. Detector simulation tasks in ATLAS are performed
via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These tasks account for
$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
offloading.

Detector simulation is part of the ATLAS production Monte Carlo (MC) workflow
(also known as MC production
chain)~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC workflow
consists of four main stages: event generation, detector simulation,
digitization, and reconstruction. Event generation creates sets of particle
four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia},
HERWIG~\cite{corcella2001herwig} and many others. \mtnote{Addresses Alexei's
comment. Is the following sentence correct?} The detector simulator is called
Geant4~\cite{agostinelli2003geant4} and simulates the ATLAS detector and the
interaction between that detector and particles.
% the interaction of these particles with the sensitive material of the ATLAS
% detector.
Each interaction creates a so-called hit and all hits are collected and passed
on for digitalization, where hits are further process to mimic the readout of
the detector. Finally, reconstruction operates local pattern recognition,
creating high-level objects like particles and jets.

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
    \caption{PanDA Broker architecture as deployed on Titan. Numbers indicates
    the execution process of a detector simulation job, part of the production
    ATLAS MD workflow.}
\label{fig:panda_broker}
\end{figure}

>>>>

Titan's architecture, configuration and policies poses several challenges to
the integration with PanDA. The \jhanote{default?} deployment model of PanDA
Pilot is unfeasible on Titan: PanDA Pilot is required to contact the Job
Dispatcher of the PanDA Server to pull jobs to execute, but this is not
possible on Titan because worker nodes do not offer outbound network
connectivity. Further, Titan does not support PanDA's security model based on
certificates and virtual organizations, making PanDA's approach to
identity management also unfeasible. While DTNs offer wide area network data
transfer, an integration with ATLAS DDM is beyond the functional and
administrative scope of the current prototyping phase. Finally, the specific
characteristics of the execution environment, especially the absence of local
storage on the worker nodes and modules tailored to Compute Node Linux,
require re-engineering of ATLAS application frameworks.

Currently, very few HEP applications can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows can
be off-loaded from the Grid to Titan's large amount of cores. Further, when HEP
tasks can be partitioned into independent jobs, Titan worker nodes can be used to
execute up to 16 concurrent payloads, one per each available core. Given these
constraints and challenges, the type of task most suitable for execution at the
moment on Titan is Monte Carlo detector simulation. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and with
small input data requirements. Detector simulation tasks in ATLAS are performed
via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These tasks account for
$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
offloading.

Detector simulation is part of the ATLAS production Monte Carlo (MC) workflow
(also known as MC production
chain)~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC workflow
consists of four main stages: event generation, detector simulation,
digitization, and reconstruction. Event generation creates sets of particle
four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia},
HERWIG~\cite{corcella2001herwig} and many others. \mtnote{Addresses Alexei's
comment. Is the following sentence correct?} The detector simulator is called
Geant4~\cite{agostinelli2003geant4} and simulates the ATLAS detector and the
interaction between that detector and particles.
% the interaction of these particles with the sensitive material of the ATLAS
% detector.
Each interaction creates a so-called hit and all hits are collected and passed
on for digitalization, where hits are further processed to mimic the readout of
the detector. Finally, reconstruction operates local pattern recognition,
creating high-level objects like particles and jets.

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
    \caption{PanDA Broker architecture as deployed on Titan. Numbers indicates
    the execution process of a detector simulation job, part of the production
    ATLAS MD workflow. \jhanote{minor error in figure: "ATLAS MC" not "MD"}}
\label{fig:panda_broker}
\end{figure}
